---
title: "DWrangling & EDA for CAT competition"
author: "Vijay Sathish"
date: "Saturday, July 04, 2015"
output: html_document
---

### Load all the libraries
```{r}

require(ggplot2)
require(GGally)
require(reshape2)
require(gridExtra)
require(dplyr)

```

### Load in all the tables

```{r}

raw.train <- read.csv("D:/Kaggle/CAT/data/train_set.csv", header = T)
raw.test <- read.csv("D:/Kaggle/CAT/data/test_set.csv", header = T)

raw.tubes <- read.csv("D:/Kaggle/CAT/data/tube.csv", header = T)
raw.bom <- read.csv("D:/Kaggle/CAT/data/bill_of_materials.csv", header = T)

raw.spec <- read.csv("D:/Kaggle/CAT/data/specs.csv", header = T)

# Probably not going to use this
raw.components <- read.csv("D:/Kaggle/CAT/data/components.csv", header = T)
raw.tube_end_form <- read.csv("D:/Kaggle/CAT/data/tube_end_form.csv", header = T)


```

Notes -
- Bracket pricing has only two levels 'yes' and 'no'
- 'id' in raw.test is a unique id for a bid; no corresponding column for the train dataset
- There are way more 'id's than 'tube_assembly_id's (4x more) which means that one type of tube gets multiple bids 
- Merge and split by 'id' wont work; we first have to generate an id for train test too
- Train and test set are both about 30k observations, so roughly equal sized

### Explore supplier stats

```{r}
common_suppliers <- intersect(levels(raw.train$supplier), levels(raw.test$supplier))
length(common_suppliers)
setdiff(levels(raw.train$supplier), levels(raw.test$supplier))
setdiff(levels(raw.test$supplier), levels(raw.train$supplier))

summary(raw.train$bracket_pricing)
summary(raw.test$bracket_pricing)

```

Notes -
- 57 suppliers in train set
- 56 suppliers in test set
- There is a substantial overlap (45 common); but it also means some suppliers in test set have no data from training set
- Note: *** Plain label encoding separately for train and test set WILL screw up the numbers. Merge suppliers to first fit label encoder, then transform separately
- Bracket pricing for ~26K points and non-bracket pricing for ~4K points in both data sets

### Tube count stats
Notes - Engineer simple counts features per tube_assembly_id
- Num specs types per id
- Num comp_id types per id
- Num comp_name types per id
- Num comp_type types per id
- Total comp_ids per id
- Total comp_name per id
- Total comp_types per id
- Note the last 3 incorporate quantities various former 3 only count unique comp_*
- Also note that last 3 are ALL the same, however you split it, so we need only one variable for those

```{r}
# 8 columns because of 5 stats plus 'tube_assembly_id'
tube_stats <- as.data.frame(matrix(0, ncol = 5, nrow = dim(raw.tubes)[1]))
names(tube_stats) <- c('uniq_specs', 'uniq_comp_ids', 'uniq_comp_types', 'uniq_comp_names', 'total_comps')
tube_stats$tube_assembly_id <- raw.tubes$tube_assembly_id


```



### Transform Bill of materials into a wide form sparse matrix - upto 8 components per 'tube_assembly_id'

```{r}

# 1. Transform each component column from long to wide form and store as a separate temporary dataframe
bom.wide1 <- dcast(raw.bom, tube_assembly_id ~ component_id_1, value.var = "quantity_1")
bom.wide1[is.na(bom.wide1)] <- 0
bom.wide2 <- dcast(raw.bom, tube_assembly_id ~ component_id_2, value.var = "quantity_2")
bom.wide2[is.na(bom.wide2)] <- 0
bom.wide3 <- dcast(raw.bom, tube_assembly_id ~ component_id_3, value.var = "quantity_3")
bom.wide3[is.na(bom.wide3)] <- 0
bom.wide4 <- dcast(raw.bom, tube_assembly_id ~ component_id_4, value.var = "quantity_4")
bom.wide4[is.na(bom.wide4)] <- 0
bom.wide5 <- dcast(raw.bom, tube_assembly_id ~ component_id_5, value.var = "quantity_5")
bom.wide5[is.na(bom.wide5)] <- 0
bom.wide6 <- dcast(raw.bom, tube_assembly_id ~ component_id_6, value.var = "quantity_6")
bom.wide6[is.na(bom.wide6)] <- 0
bom.wide7 <- dcast(raw.bom, tube_assembly_id ~ component_id_7, value.var = "quantity_7")
bom.wide7[is.na(bom.wide7)] <- 0
bom.wide8 <- dcast(raw.bom, tube_assembly_id ~ component_id_8, value.var = "quantity_8")
bom.wide8[is.na(bom.wide8)] <- 0

# 1a. Get tube_stats from BOM
# Convert the NAs to 0 prior to summing
raw.bom[is.na(raw.bom)] <- 0
# Count tot_comps per tube_assembly_id
tube_stats$total_comps = raw.bom$quantity_1 + raw.bom$quantity_2 +
    raw.bom$quantity_3 + raw.bom$quantity_4 +
    raw.bom$quantity_5 + raw.bom$quantity_6 +
    raw.bom$quantity_7 + raw.bom$quantity_8
# Count uniq_comp_ids per tube_assembly_id
tube_stats$uniq_comp_ids = as.numeric(!is.na(raw.bom$component_id_1)) + as.numeric(!is.na(raw.bom$component_id_2)) +
    as.numeric(!is.na(raw.bom$component_id_3)) + as.numeric(!is.na(raw.bom$component_id_4)) +
    as.numeric(!is.na(raw.bom$component_id_5)) + as.numeric(!is.na(raw.bom$component_id_6)) +
    as.numeric(!is.na(raw.bom$component_id_7)) + as.numeric(!is.na(raw.bom$component_id_8))

# 2. Get the union of component names C-xxxx
components <- union(names(bom.wide1), names(bom.wide2))
components <- union(components, names(bom.wide3))
components <- union(components, names(bom.wide4))
components <- union(components, names(bom.wide5))
components <- union(components, names(bom.wide6))
components <- union(components, names(bom.wide7))
components <- union(components, names(bom.wide8))
length(components)

# 3. Create an empty output dataframe with correct dimensions and column names
bom.wide <- as.data.frame(matrix(0, ncol = length(components), nrow = dim(bom.wide1)[1]))
names(bom.wide) <- components
# 3a. Initialize the tube_assembly_id for bom.wide
bom.wide$tube_assembly_id <- bom.wide1$tube_assembly_id

# 4. Remove "tube_assembly_id" from components
components <- setdiff(components, "tube_assembly_id")

# 5. Combine all the temporary data frames into a single wide dataframe by summing up component quantities
# http://stackoverflow.com/questions/19730806/r-access-data-frame-column-using-variable
for (col in components) {
  if (length(intersect(names(bom.wide1), col))) {
    bom.wide[, col] <- bom.wide[, col] + bom.wide1[, col]
  }
  if (length(intersect(names(bom.wide2), col))) {
    bom.wide[, col] <- bom.wide[, col] + bom.wide2[, col]
  }
  if (length(intersect(names(bom.wide3), col))) {
    bom.wide[, col] <- bom.wide[, col] + bom.wide3[, col]
  }
  if (length(intersect(names(bom.wide4), col))) {
    bom.wide[, col] <- bom.wide[, col] + bom.wide4[, col]
  }
  if (length(intersect(names(bom.wide5), col))) {
    bom.wide[, col] <- bom.wide[, col] + bom.wide5[, col]
  }
  if (length(intersect(names(bom.wide6), col))) {
    bom.wide[, col] <- bom.wide[, col] + bom.wide6[, col]
  }
  if (length(intersect(names(bom.wide7), col))) {
    bom.wide[, col] <- bom.wide[, col] + bom.wide7[, col]
  }
  if (length(intersect(names(bom.wide8), col))) {
    bom.wide[, col] <- bom.wide[, col] + bom.wide8[, col]
  }
}

# 5a. Get rid of the 'NA' column (normally, empty value is interpreted as na, but since csv comes with 'NA', get rid of that)
bom.wide$"NA" <- NULL
dim(bom.wide)

# 6. Remove all temporary dataframes
remove(bom.wide1)
remove(bom.wide2)
remove(bom.wide3)
remove(bom.wide4)
remove(bom.wide5)
remove(bom.wide6)
remove(bom.wide7)
remove(bom.wide8)

# 7. Write the transformed output back to CSV
write.csv(bom.wide, "D:/Kaggle/CAT/data/transformed/bom_wide.csv", row.names = F, na = "", sep = ",", col.names = T)

```

Notes -
- *** Total of 2048 components after the wide form transformation
- Since, dealing with wide form like this will be difficult, apply TfidfTransformer followed by TruncatedSVD to reduce feature set from BOM

### Remap components to component_types by joining bill_of_materials.csv and components.csv
```{r}

# Go through each component variable in raw.bom and re-map to component type based on components.csv

map_comp_to_comp_type <- function(df_new, x_join_var, new_var_name) {
  #x_join_var = 'component_id_1'
  #new_var_name = 'comp_type_1'
  # Go through each component variable in raw.bom and assign quantity to component type based on components.csv
  df_new <- merge(x = df_new, y = raw.components, by.x = x_join_var, by.y = 'component_id', all.x = T, incomparables = NA)
  # Rename the 'component_type_id' column and drop 'name' and 'component_id_x' column after join
  names(df_new)[names(df_new) == "component_type_id"] <- new_var_name
  df_new$name <- NULL
  df_new$x_join_var <- NULL
  return (df_new)
  
}

raw.comp_type <- raw.bom
raw.comp_type = map_comp_to_comp_type(raw.comp_type, 'component_id_1', 'comp_type_1')
raw.comp_type = map_comp_to_comp_type(raw.comp_type, 'component_id_2', 'comp_type_2')
raw.comp_type = map_comp_to_comp_type(raw.comp_type, 'component_id_3', 'comp_type_3')
raw.comp_type = map_comp_to_comp_type(raw.comp_type, 'component_id_4', 'comp_type_4')
raw.comp_type = map_comp_to_comp_type(raw.comp_type, 'component_id_5', 'comp_type_5')
raw.comp_type = map_comp_to_comp_type(raw.comp_type, 'component_id_6', 'comp_type_6')
raw.comp_type = map_comp_to_comp_type(raw.comp_type, 'component_id_7', 'comp_type_7')
raw.comp_type = map_comp_to_comp_type(raw.comp_type, 'component_id_8', 'comp_type_8')
# Sort dataframe by 'tube_assembly_id'
raw.comp_type <- arrange(raw.comp_type, tube_assembly_id)

```

Notes - 
- 29 component types in all; hence much more compressed format than component_id


### Transform newly minted raw.comp_type to wide form sparse matrix
```{r}
# 1. Transform each component column from long to wide form and store as a separate temporary dataframe
comp_type.wide1 <- dcast(raw.comp_type, tube_assembly_id ~ comp_type_1, value.var = "quantity_1")
comp_type.wide1[is.na(comp_type.wide1)] <- 0
comp_type.wide2 <- dcast(raw.comp_type, tube_assembly_id ~ comp_type_2, value.var = "quantity_2")
comp_type.wide2[is.na(comp_type.wide2)] <- 0
comp_type.wide3 <- dcast(raw.comp_type, tube_assembly_id ~ comp_type_3, value.var = "quantity_3")
comp_type.wide3[is.na(comp_type.wide3)] <- 0
comp_type.wide4 <- dcast(raw.comp_type, tube_assembly_id ~ comp_type_4, value.var = "quantity_4")
comp_type.wide4[is.na(comp_type.wide4)] <- 0
comp_type.wide5 <- dcast(raw.comp_type, tube_assembly_id ~ comp_type_5, value.var = "quantity_5")
comp_type.wide5[is.na(comp_type.wide5)] <- 0
comp_type.wide6 <- dcast(raw.comp_type, tube_assembly_id ~ comp_type_6, value.var = "quantity_6")
comp_type.wide6[is.na(comp_type.wide6)] <- 0
comp_type.wide7 <- dcast(raw.comp_type, tube_assembly_id ~ comp_type_7, value.var = "quantity_7")
comp_type.wide7[is.na(comp_type.wide7)] <- 0
comp_type.wide8 <- dcast(raw.comp_type, tube_assembly_id ~ comp_type_8, value.var = "quantity_8")
comp_type.wide8[is.na(comp_type.wide8)] <- 0

# 2. Get the union of component names C-xxxx
num_comp_types <- length(levels(raw.components$component_type_id))
comp_type_names <- levels(raw.components$component_type_id)

# 3. Create an empty output dataframe with correct dimensions and column names
comp_type.wide <- as.data.frame(matrix(0, ncol = num_comp_types, nrow = dim(bom.wide)[1]))
names(comp_type.wide) <- comp_type_names

# 4. Initialize the tube_assembly_id for comp_type.wide
comp_type.wide$tube_assembly_id <- comp_type.wide1$tube_assembly_id

# 5. Combine all the temporary data frames into a single wide dataframe by summing up component quantities
# http://stackoverflow.com/questions/19730806/r-access-data-frame-column-using-variable
for (col in comp_type_names) {
  if (length(intersect(names(comp_type.wide1), col))) {
    comp_type.wide[, col] <- comp_type.wide[, col] + comp_type.wide1[, col]
  }
  if (length(intersect(names(comp_type.wide2), col))) {
    comp_type.wide[, col] <- comp_type.wide[, col] + comp_type.wide2[, col]
  }
  if (length(intersect(names(comp_type.wide3), col))) {
    comp_type.wide[, col] <- comp_type.wide[, col] + comp_type.wide3[, col]
  }
  if (length(intersect(names(comp_type.wide4), col))) {
    comp_type.wide[, col] <- comp_type.wide[, col] + comp_type.wide4[, col]
  }
  if (length(intersect(names(comp_type.wide5), col))) {
    comp_type.wide[, col] <- comp_type.wide[, col] + comp_type.wide5[, col]
  }
  if (length(intersect(names(comp_type.wide6), col))) {
    comp_type.wide[, col] <- comp_type.wide[, col] + comp_type.wide6[, col]
  }
  if (length(intersect(names(comp_type.wide7), col))) {
    comp_type.wide[, col] <- comp_type.wide[, col] + comp_type.wide7[, col]
  }
  if (length(intersect(names(comp_type.wide8), col))) {
    comp_type.wide[, col] <- comp_type.wide[, col] + comp_type.wide8[, col]
  }
}

# 5a. Get rid of the 'NA' column (normally, empty value is interpreted as na, but since csv comes with 'NA', get rid of that)
comp_type.wide$"NA" <- NULL
dim(comp_type.wide)

# 6. Remove all temporary dataframes
remove(comp_type.wide1)
remove(comp_type.wide2)
remove(comp_type.wide3)
remove(comp_type.wide4)
remove(comp_type.wide5)
remove(comp_type.wide6)
remove(comp_type.wide7)
remove(comp_type.wide8)

# 7. Write the transformed output back to CSV
write.csv(comp_type.wide, "D:/Kaggle/CAT/data/transformed/comp_type_wide.csv", row.names = F, na = "", sep = ",", col.names = T)

# 8 Get tube_stats from comp_type
# Count uniq_comp_types per tube_assembly_id
accumulate_unique_comp_types <- function(df_new,  col_name) {
  df_new$uniq_comp_types <- df_new$uniq_comp_types + as.numeric(comp_type.wide[, col_name] > 0)
  return (df_new)
}

# Iterate over all comp_type columns
for (col in comp_type_names) {
  tube_stats <- accumulate_unique_comp_types(tube_stats, col)
}


```

### Go through each component variable in raw.bom and re-map to component name based on components.csv
```{r}
map_comp_to_comp_name <- function(df_new, x_join_var, new_var_name) {
  #x_join_var = 'component_id_1'
  #new_var_name = 'comp_name_1'
  # Go through each component variable in raw.bom and assign quantity to component name based on components.csv
  df_new <- merge(x = df_new, y = raw.components, by.x = x_join_var, by.y = 'component_id', all.x = T, incomparables = NA)
  # Rename the 'name' column and drop 'component_type_id' and 'component_id_x' column after join
  names(df_new)[names(df_new) == "name"] <- new_var_name
  df_new$component_type_id <- NULL
  df_new$x_join_var <- NULL
  return (df_new)
  
}

raw.comp_name <- raw.bom
raw.comp_name = map_comp_to_comp_name(raw.comp_name, 'component_id_1', 'comp_name_1')
raw.comp_name = map_comp_to_comp_name(raw.comp_name, 'component_id_2', 'comp_name_2')
raw.comp_name = map_comp_to_comp_name(raw.comp_name, 'component_id_3', 'comp_name_3')
raw.comp_name = map_comp_to_comp_name(raw.comp_name, 'component_id_4', 'comp_name_4')
raw.comp_name = map_comp_to_comp_name(raw.comp_name, 'component_id_5', 'comp_name_5')
raw.comp_name = map_comp_to_comp_name(raw.comp_name, 'component_id_6', 'comp_name_6')
raw.comp_name = map_comp_to_comp_name(raw.comp_name, 'component_id_7', 'comp_name_7')
raw.comp_name = map_comp_to_comp_name(raw.comp_name, 'component_id_8', 'comp_name_8')
# Sort dataframe by 'tube_assembly_id'
raw.comp_name <- arrange(raw.comp_name, tube_assembly_id)

```

Notes - 
- 297 component names in all; still a much more compressed format than component_id but 10x more categories than component_types


### Transform newly minted raw.comp_name to wide form sparse matrix
```{r}
# 1. Transform each component column from long to wide form and store as a separate temporary dataframe
comp_name.wide1 <- dcast(raw.comp_name, tube_assembly_id ~ comp_name_1, value.var = "quantity_1")
comp_name.wide1[is.na(comp_name.wide1)] <- 0
comp_name.wide2 <- dcast(raw.comp_name, tube_assembly_id ~ comp_name_2, value.var = "quantity_2")
comp_name.wide2[is.na(comp_name.wide2)] <- 0
comp_name.wide3 <- dcast(raw.comp_name, tube_assembly_id ~ comp_name_3, value.var = "quantity_3")
comp_name.wide3[is.na(comp_name.wide3)] <- 0
comp_name.wide4 <- dcast(raw.comp_name, tube_assembly_id ~ comp_name_4, value.var = "quantity_4")
comp_name.wide4[is.na(comp_name.wide4)] <- 0
comp_name.wide5 <- dcast(raw.comp_name, tube_assembly_id ~ comp_name_5, value.var = "quantity_5")
comp_name.wide5[is.na(comp_name.wide5)] <- 0
comp_name.wide6 <- dcast(raw.comp_name, tube_assembly_id ~ comp_name_6, value.var = "quantity_6")
comp_name.wide6[is.na(comp_name.wide6)] <- 0
comp_name.wide7 <- dcast(raw.comp_name, tube_assembly_id ~ comp_name_7, value.var = "quantity_7")
comp_name.wide7[is.na(comp_name.wide7)] <- 0
comp_name.wide8 <- dcast(raw.comp_name, tube_assembly_id ~ comp_name_8, value.var = "quantity_8")
comp_name.wide8[is.na(comp_name.wide8)] <- 0

# 2. Get the union of component names C-xxxx
num_comp_names <- length(levels(raw.components$name))
comp_name_names <- levels(raw.components$name)

# 3. Create an empty output dataframe with correct dimensions and column names
comp_name.wide <- as.data.frame(matrix(0, ncol = num_comp_names, nrow = dim(bom.wide)[1]))
names(comp_name.wide) <- comp_name_names

# 4. Initialize the tube_assembly_id for comp_name.wide
comp_name.wide$tube_assembly_id <- comp_name.wide1$tube_assembly_id

# 5. Combine all the temporary data frames into a single wide dataframe by summing up component quantities
# http://stackoverflow.com/questions/19730806/r-access-data-frame-column-using-variable
for (col in comp_name_names) {
  if (length(intersect(names(comp_name.wide1), col))) {
    comp_name.wide[, col] <- comp_name.wide[, col] + comp_name.wide1[, col]
  }
  if (length(intersect(names(comp_name.wide2), col))) {
    comp_name.wide[, col] <- comp_name.wide[, col] + comp_name.wide2[, col]
  }
  if (length(intersect(names(comp_name.wide3), col))) {
    comp_name.wide[, col] <- comp_name.wide[, col] + comp_name.wide3[, col]
  }
  if (length(intersect(names(comp_name.wide4), col))) {
    comp_name.wide[, col] <- comp_name.wide[, col] + comp_name.wide4[, col]
  }
  if (length(intersect(names(comp_name.wide5), col))) {
    comp_name.wide[, col] <- comp_name.wide[, col] + comp_name.wide5[, col]
  }
  if (length(intersect(names(comp_name.wide6), col))) {
    comp_name.wide[, col] <- comp_name.wide[, col] + comp_name.wide6[, col]
  }
  if (length(intersect(names(comp_name.wide7), col))) {
    comp_name.wide[, col] <- comp_name.wide[, col] + comp_name.wide7[, col]
  }
  if (length(intersect(names(comp_name.wide8), col))) {
    comp_name.wide[, col] <- comp_name.wide[, col] + comp_name.wide8[, col]
  }
}

# 5a. Get rid of the 'NA' column (normally, empty value is interpreted as na, but since csv comes with 'NA', get rid of that)
comp_name.wide$"NA" <- NULL
dim(comp_name.wide)

# 6. Remove all temporary dataframes
remove(comp_name.wide1)
remove(comp_name.wide2)
remove(comp_name.wide3)
remove(comp_name.wide4)
remove(comp_name.wide5)
remove(comp_name.wide6)
remove(comp_name.wide7)
remove(comp_name.wide8)

# 7. Write the transformed output back to CSV
write.csv(comp_name.wide, "D:/Kaggle/CAT/data/transformed/comp_name_wide.csv", row.names = F, na = "", sep = ",", col.names = T)

# 8 Get tube_stats from comp_name
# Count uniq_comp_names per tube_assembly_id
accumulate_unique_comp_names <- function(df_new,  col_name) {
  df_new$uniq_comp_names <- df_new$uniq_comp_names + as.numeric(comp_name.wide[, col_name] > 0)
  return (df_new)
}

# Iterate over all comp_name columns
for (col in comp_name_names) {
  tube_stats <- accumulate_unique_comp_names(tube_stats, col)
}

```


### Transform Specs into a wide form sparse matrix - upto 10 specs per 'tube_assembly_id'

```{r}

# 1. Transform each component column from long to wide form and store as a separate temporary dataframe
raw.spec$specq_1 <- 1
spec.wide1 <- dcast(raw.spec, tube_assembly_id ~ spec1, value.var = "specq_1")
spec.wide1[is.na(spec.wide1)] <- 0
raw.spec$specq_2 <- 1
spec.wide2 <- dcast(raw.spec, tube_assembly_id ~ spec2, value.var = "specq_2")
spec.wide2[is.na(spec.wide2)] <- 0
raw.spec$specq_3 <- 1
spec.wide3 <- dcast(raw.spec, tube_assembly_id ~ spec3, value.var = "specq_3")
spec.wide3[is.na(spec.wide3)] <- 0
raw.spec$specq_4 <- 1
spec.wide4 <- dcast(raw.spec, tube_assembly_id ~ spec4, value.var = "specq_4")
spec.wide4[is.na(spec.wide4)] <- 0
raw.spec$specq_5 <- 1
spec.wide5 <- dcast(raw.spec, tube_assembly_id ~ spec5, value.var = "specq_5")
spec.wide5[is.na(spec.wide5)] <- 0
raw.spec$specq_6 <- 1
spec.wide6 <- dcast(raw.spec, tube_assembly_id ~ spec6, value.var = "specq_6")
spec.wide6[is.na(spec.wide6)] <- 0
raw.spec$specq_7 <- 1
spec.wide7 <- dcast(raw.spec, tube_assembly_id ~ spec7, value.var = "specq_7")
spec.wide7[is.na(spec.wide7)] <- 0
raw.spec$specq_8 <- 1
spec.wide8 <- dcast(raw.spec, tube_assembly_id ~ spec8, value.var = "specq_8")
spec.wide8[is.na(spec.wide8)] <- 0
raw.spec$specq_9 <- 1
spec.wide9 <- dcast(raw.spec, tube_assembly_id ~ spec9, value.var = "specq_9")
spec.wide9[is.na(spec.wide9)] <- 0
raw.spec$specq_10 <- 1
spec.wide10 <- dcast(raw.spec, tube_assembly_id ~ spec10, value.var = "specq_10")
spec.wide10[is.na(spec.wide10)] <- 0
  
# 1a. Get tube_stats from specifications
# Convert the NAs to 0 prior to summing
raw.spec[is.na(raw.spec)] <- 0
# Count uniq_specs per tube_assembly_id
tube_stats$uniq_specs = as.numeric(!is.na(raw.spec$spec1)) + as.numeric(!is.na(raw.spec$spec2)) +
    as.numeric(!is.na(raw.spec$spec3)) + as.numeric(!is.na(raw.spec$spec4)) +
    as.numeric(!is.na(raw.spec$spec5)) + as.numeric(!is.na(raw.spec$spec6)) +
    as.numeric(!is.na(raw.spec$spec7)) + as.numeric(!is.na(raw.spec$spec8)) +
    as.numeric(!is.na(raw.spec$spec9)) + as.numeric(!is.na(raw.spec$spec10))


# 2. Get the union of spec names SP-xxxx
specs <- union(names(spec.wide1), names(spec.wide2))
specs <- union(specs, names(spec.wide3))
specs <- union(specs, names(spec.wide4))
specs <- union(specs, names(spec.wide5))
specs <- union(specs, names(spec.wide6))
specs <- union(specs, names(spec.wide7))
specs <- union(specs, names(spec.wide8))
specs <- union(specs, names(spec.wide9))
specs <- union(specs, names(spec.wide10))
length(specs)

# 3. Create an empty output dataframe with correct dimensions and column names
spec.wide <- as.data.frame(matrix(0, ncol = length(specs), nrow = dim(spec.wide1)[1]))
names(spec.wide) <- specs
# 3a. Initialize the tube_assembly_id for spec.wide
spec.wide$tube_assembly_id <- spec.wide1$tube_assembly_id

# 4. Remove "tube_assembly_id" from specs
specs <- setdiff(specs, "tube_assembly_id")

# 5. Combine all the temporary data frames into a single wide dataframe by summing up spec quantities
# http://stackoverflow.com/questions/19730806/r-access-data-frame-column-using-variable
for (col in specs) {
  if (length(intersect(names(spec.wide1), col))) {
    spec.wide[, col] <- spec.wide[, col] + spec.wide1[, col]
  }
  if (length(intersect(names(spec.wide2), col))) {
    spec.wide[, col] <- spec.wide[, col] + spec.wide2[, col]
  }
  if (length(intersect(names(spec.wide3), col))) {
    spec.wide[, col] <- spec.wide[, col] + spec.wide3[, col]
  }
  if (length(intersect(names(spec.wide4), col))) {
    spec.wide[, col] <- spec.wide[, col] + spec.wide4[, col]
  }
  if (length(intersect(names(spec.wide5), col))) {
    spec.wide[, col] <- spec.wide[, col] + spec.wide5[, col]
  }
  if (length(intersect(names(spec.wide6), col))) {
    spec.wide[, col] <- spec.wide[, col] + spec.wide6[, col]
  }
  if (length(intersect(names(spec.wide7), col))) {
    spec.wide[, col] <- spec.wide[, col] + spec.wide7[, col]
  }
  if (length(intersect(names(spec.wide8), col))) {
    spec.wide[, col] <- spec.wide[, col] + spec.wide8[, col]
  }
  if (length(intersect(names(spec.wide9), col))) {
    spec.wide[, col] <- spec.wide[, col] + spec.wide9[, col]
  }
  if (length(intersect(names(spec.wide10), col))) {
    spec.wide[, col] <- spec.wide[, col] + spec.wide10[, col]
  }
}

# 5a. Get rid of the 'NA' column (normally, empty value is interpreted as na, but since csv comes with 'NA', get rid of that)
spec.wide$"NA" <- NULL
dim(spec.wide)

# 6. Remove all temporary dataframes
remove(spec.wide1)
remove(spec.wide2)
remove(spec.wide3)
remove(spec.wide4)
remove(spec.wide5)
remove(spec.wide6)
remove(spec.wide7)
remove(spec.wide8)
remove(spec.wide9)
remove(spec.wide10)

# 7. Write the transformed output back to CSV
write.csv(spec.wide, "D:/Kaggle/CAT/data/transformed/spec_wide.csv", row.names = F, na = "", sep = ",", col.names = T)

```
- *** Total of 85 specs after the wide form transformation
- Number of specs are limited compared to BOM components, so maybe we don't TfidfTransformer + SVD; need to see

### Merge tubes.csv and tube_end_form.csv and only retain rows from left table
```{r}

tubes <- merge(x = raw.tubes, y = raw.tube_end_form, by.x = "end_a", by.y = "end_form_id", all.x = T)
tubes$a_forming <- tubes$forming
tubes$forming <- NULL
tubes <- merge(x = tubes, y = raw.tube_end_form, by.x = "end_x", by.y = "end_form_id", all.x = T)
tubes$x_forming <- tubes$forming
tubes$forming <- NULL

# Convert all Factor Variables to numeric to avoid label encoding again in Python
# Binary variables
tubes$x_forming <- as.numeric(tubes$x_forming) - 1
tubes$a_forming <- as.numeric(tubes$a_forming) - 1
tubes$end_a_1x <- as.numeric(tubes$end_a_1x) - 1
tubes$end_a_2x <- as.numeric(tubes$end_a_2x) - 1
tubes$end_x_1x <- as.numeric(tubes$end_x_1x) - 1
tubes$end_x_2x <- as.numeric(tubes$end_x_2x) - 1

# Factored variables
tubes$end_x <- as.numeric(tubes$end_x) - 1
tubes$end_a <- as.numeric(tubes$end_a) - 1
tubes$material_id <- as.numeric(tubes$material_id) - 1

# Calculate the amount of tube material
tubes$tube_material = with(tubes, (pi*(diameter/2)^2 - pi*((diameter - wall)/2)^2)*length)

# Write transformed tubes back as CSV (Write NA as empty value so that Python can pick it up correctly)
write.csv(tubes, "D:/Kaggle/CAT/data/transformed/tubes_transformed.csv", row.names = F, na = "", sep = ",", col.names = T)

# Create another with tube_stats (retain older one as well)
tubes2 <- merge(x = tubes, y = tube_stats, by = 'tube_assembly_id', all.x = T)

# Write transformed tubes2 back as CSV (Write NA as empty value so that Python can pick it up correctly)
write.csv(tubes2, "D:/Kaggle/CAT/data/transformed/tubes_transformed_v2.csv", row.names = F, na = "", sep = ",", col.names = T)


```


